Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-09-07 manuscript/response_to_reviews.Rmd

Other comments:

- As mentioned on a previous comment, in the manuscript the authors sampled dead and live trees, but then I did not see any further mentioning to this later in the manuscript (lines 175-178 and Table 2). This would merit some discussion and further analysis, how many trees were died and how many trees were alive of those in Table 2?

*I think our argument is pretty strong, but we need to make some change to respond to this comment. Looking at resilience and recovery should help; presuming that they are high, we have a further argument that death was not caused by drought, and hence that status at coring doesn't matter.*

**We consider it extremely unlikely that the status at coring would affect drought response. We have modified the text at (former) lines 175-178 to make this more clear: "We note that drought was probably not a cause of mortality for these trees, as monthly May-Aug $PDSI$ did not drop below -1.75 in these years or the three years prior (2013-2017), and that trees cored dead displayed similar climate sensitivity to trees cored live [@helcoski_growing_2019]. Lagged drought-induced mortality would be unlikely, given that the trees analyzed here lived at least 17-18 years past the most recent major drought (1999), whereas the meta-analysis of @trugman_tree_2018 indicates that >10-year lags in drought-attributed mortality are rare."**

**We have added a footnote to Table 2 indicating where the information on numbers cored live vs dead may be found (Table S1 of Helcoski et al. 2019).**

@@

**We now define AICc at first mention, which is in the paragraph above. The sentence now reads "We used Akaike information criterion with correction for small sample sizes (AICc) to assess model selection,..." In the sentence referenced here, we define $\Delta$AIC: "Trait variables were considered appropriate for inclusion in the main model if they had a consistent direction of response across all droughts and if their addition to the base model improved fit (at $\Delta$AICc $\ge$ 1.0, where $\Delta$AICc is the difference in AICc between models with and without the trait) in at least one drought year (Table S4)."**

**We have given careful thought to the trait variable selection criteria. It is important to note that selection of traits is not a test of their significance in the final model; it is simply a way of objectively reducing the number of candidate traits in these models. While we fully agree that $\Delta$AICc $\ge$ 2 is the appropriate criterion for assessing significance, it is not necessarily appropriate for candidate variable selection because variable significance changes when considered independently versus in the context of a full model. That is, the significance of variables--particularly those with marginal significance-- cannot be reliably assessed outside the context of a model that simultaneously accounts for other relevant variables. As a case in point, $\pi_{tlp}$ would be removed by a $\Delta$AICc $\ge$ 2 criterion for variable selection, but is included in the best model for all droughts combined, as well as the 1977 drought. While it's effects are marginal (i.e., included in some but not all top models), we feel that it is important to include it as a candidate variable both for the statistical considerations mentioned above, and because we expect that a lot of readers will be very interested to see a full assessment of this trait, which is currently of great interest to the research community.**

*([issue #134](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/134))*

### 2020-09-07 manuscript/climate_sensitivity_draft.Rmd
more on R3 comment on live/dead

Text | word count | other | n
---------------------------------------------|---------------------|-----------------------------|--------------------
Total word count (excluding summary, references and legends) | 6,490	 |	No. of figures | 	4 (all colour)
Summary|  	198 | 	No. of Tables | 	3  
Introduction|	1,264 |	No of Supporting Information files | 	14
Materials and Methods |	2,228	||	 
Results |	929	 ||
Discussion |	1927	||
Acknowledgements | 142	||

@@

We analyzed tree-ring data (xylem growth increment) from 571 trees representing the twelve dominant species (Table 2; Fig. **S2**). 
Selected species were those with the greatest contributions to woody aboveground net primary productivity ($ANPP_{stem}$) and together comprised 97% of study plot $ANPP_{stem}$ between 2008 and 2013 [@helcoski_growing_2019]. 
Cores (one per tree) were collected within the ForestGEO plot at breast height (1.3m) in 2010-2011 or 2016-2017. 
In 2010-2011, cores were collected from randomly selected live trees of each species that had at least 30 individuals $\ge$ 10 cm DBH [@bourg_initial_2013]. 
Annual tree mortality censuses were initiated in 2014 [@gonzalezakre_patterns_2016], and in 2016-2017, cores were collected from all trees found to have died since the previous year’s census. 
We note that drought was probably not a cause of mortality for these trees, as monthly May-Aug $PDSI$ did not drop below -1.75 in these years or the three years prior (2013-2017), and that trees cored dead displayed similar climate sensitivity to trees cored live [@helcoski_growing_2019]. 
Cores were sanded, measured, and crossdated using standard procedures, as detailed in [@helcoski_growing_2019]. 
The resulting chronologies (Fig. **1a**) were published in Zenodo (DOI: 10.5281/zenodo.2649302) in association with @helcoski_growing_2019.

>>>>

Text | word count | other | n
---------------------------------------------|---------------------|-----------------------------|--------------------
Total word count (excluding summary, references and legends) | 6,490	 |	No. of figures | 	4 (all colour)
Summary|  	198 | 	No. of Tables | 	3  
Introduction|	1,264 |	No of Supporting Information files | 	14
Materials and Methods |	2,228	||	 
Results |	929	 ||
Discussion |	1927	||
Acknowledgments | 142	||

@@

We analyzed tree-ring data (xylem growth increment) from 571 trees representing the twelve dominant species (Table 2; Fig. **S2**). 
Selected species were those with the greatest contributions to woody aboveground net primary productivity ($ANPP_{stem}$) and together comprised 97% of study plot $ANPP_{stem}$ between 2008 and 2013 [@helcoski_growing_2019]. 
Cores (one per tree) were collected within the ForestGEO plot at breast height (1.3m) in 2010-2011 or 2016-2017. 
In 2010-2011, cores were collected from randomly selected live trees of each species that had at least 30 individuals $\ge$ 10 cm DBH [@bourg_initial_2013]. 
Annual tree mortality censuses were initiated in 2014 [@gonzalezakre_patterns_2016], and in 2016-2017, cores were collected from all trees found to have died since the previous year’s census. 
We note that drought was probably not a cause of mortality for these trees, as monthly May-Aug $PDSI$ did not drop below -1.75 in these years or the three years prior (2013-2017), and that trees cored dead displayed similar climate sensitivity to trees cored live [@helcoski_growing_2019]. 
Lagged drought-induced mortality would be unlikely, given that the trees analyzed here lived at least 17-18 years past the most recent major drought (1999), whereas the meta-analysis of @trugman_tree_2018 indicates that >10-year lags in drought-attributed mortality are rare.
Cores were sanded, measured, and crossdated using standard procedures, as detailed in [@helcoski_growing_2019]. 
The resulting chronologies (Fig. **1a**) were published in Zenodo (DOI: 10.5281/zenodo.2649302) in association with @helcoski_growing_2019.

### 2020-09-07 manuscript/tables&figures.Rmd
more on R3 comment on live/dead

```{r, Table 2, eval=TRUE, echo=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
table2 <- read.csv("tables_figures/publication/table2_species_table_hand_formatted.csv", stringsAsFactors = FALSE, check.names = FALSE)
table2[8,8] <- paste0(table2[8,8], footnote_marker_symbol(1))
kable(table2, format = "latex", booktabs=TRUE, caption = "Table 2. Overview of analyzed species, listed in order of their relative contributions to woody stem productivity ($ANPP_{stem}$) in the plot, along with numbers and sizes sampled, and species traits. Variable abbreviations are as in Table 3. DBH measurements are from the most recent ForestGEO census in 2018 (live trees) or tree mortality censuses in 2016 and 2017 (trees cored dead).", escape=FALSE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 8.5, protect_latex = TRUE) %>%
  add_header_above(c(" "," "," ", "contemporary DBH (cm)" = 2, "species traits (mean +/- se)" = 5)) %>%
  #kable_styling(bootstrap_options = "striped", full_width = FALSE)
kableExtra::footnote(symbol = c("Semi-ring porosity is intermediate between ring and diffuse. We group it with diffuse-porous species for more even division of species between categories."), footnote_as_chunk = FALSE)
```

>>>>

```{r, Table 2, eval=TRUE, echo=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
table2 <- read.csv("tables_figures/publication/table2_species_table_hand_formatted.csv", stringsAsFactors = FALSE, check.names = FALSE)
table2[0,3] <- paste0(table2[0,3], footnote_marker_symbol(1))
table2[8,8] <- paste0(table2[8,8], footnote_marker_symbol(2))
kable(table2, format = "latex", booktabs=TRUE, caption = "Table 2. Overview of analyzed species, listed in order of their relative contributions to woody stem productivity ($ANPP_{stem}$) in the plot, along with numbers and sizes sampled, and species traits. Variable abbreviations are as in Table 3. DBH measurements are from the most recent ForestGEO census in 2018 (live trees) or tree mortality censuses in 2016 and 2017 (trees cored dead).", escape=FALSE) %>%
  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 8.5, protect_latex = TRUE) %>%
  add_header_above(c(" "," "," ", "contemporary DBH (cm)" = 2, "species traits (mean +/- se)" = 5)) %>%
  #kable_styling(bootstrap_options = "striped", full_width = FALSE)
kableExtra::footnote(symbol = c("Numbers cored live versus dead are given in Table S1 of Helcoski et al. (2019).", "Semi-ring porosity is intermediate between ring and diffuse. We group it with diffuse-porous species for more even division of species between categories."), footnote_as_chunk = FALSE)
```
